---
title: "Configure project and RevBayes"
author: Friederike Wölke
date: 2026-06-20
format: html
embed-resources: true
standalone: true
toc: true

---
General setup for RevBayes via Positron Interface and R.


- I use positron because it was created for R but is based on VSCode and thus allows me to use bash, python and R in the same workspace.
- I use bash as terminal language in a linux subsytem to prepare the data (and move it from my linux subsystem to windows where this workspace is located)
- I use bash also to start RevBayes and source .Rev scripts.


# I. Create project folder

```{r}
#| eval: false
# Set up workspace
dir.create("RevBayes-DEC-Uapaca", showWarnings = FALSE)
setwd("RevBayes-DEC-Uapaca")

# create content folder
dir.create("content", showWarnings = FALSE)

# create scripts folder
dir.create("scripts", showWarnings = FALSE)
```

# II. Set up R environment

```{r}
#| eval: false
packages_to_load <- c(
  "renv",
  "here",
  "ape",
  "dplyr",
  "RevGadgets",
  "magick",
  "ggplot2",
  "BiocManager"
)

lapply(packages_to_load, function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
  }
})

BiocManager::install("ggtree")

# activate renv workspace
renv::activate()

# renv::status()
# renv::install()
# renv::snapshot()
```

```{r}
# vector with packages
load_pckgs <-
  c("here", "ape", "dplyr")

# activate them
lapply(
  load_pckgs,
  library,
  character.only = TRUE
)
```


# III. Prepare input data for RevBayes
## a) nuclear tree
```{r}
#| eval: false
# the tree (output from nuclear phylogenomics pipeline)
tr_nuc <-
  read.tree(
    here(
      "content/data/input/phylogenetic_tree/wASTRAL_HybSuite_HRS.rr.tre"
    )
  ) %>%
  ladderize()

plot(tr_nuc)

# checks on the tree:
is.rooted(tr_nuc) # yes
is.binary(tr_nuc) # yes
is.ultrametric(tr_nuc) # no

summary(tr_nuc$edge.length)
tr_nuc$edge.length[tr_nuc$edge.length <= 1e-10]


# create ultrametic starting tree:
tr_nuc <- multi2di(tr_nuc)
tr_nuc$edge.length[is.na(tr_nuc$edge.length) | tr_nuc$edge.length <= 0] <- 1e-6

tr_nuc_ultra <- chronos(tr_nuc)
tr_nuc_ultra <- compute.brlen(tr_nuc_ultra, method = "Grafen")

write.tree(
  tr_nuc_ultra,
  here(
    "content/data/input/phylogenetic_tree/uapaca_start_ultrametric.tre"
  )
)
```

## b) plastome tree
```{r}
#| eval: false
tr_plastome <-
  read.tree(
    here(
      "content/data/input/phylogenetic_tree/plastome.fasta.concat.treefile"
    )
  ) %>%
  root(outgroup = "Spondianthus_preussii", resolve.root = TRUE) %>%
  ladderize()


plot(tr_plastome)


# checks on the tree:
is.rooted(tr_plastome) # yes
is.binary(tr_plastome) # yes
is.ultrametric(tr_plastome) # no

summary(tr_plastome$edge.length)
tr_plastome$edge.length[tr_plastome$edge.length <= 1e-10]


# create ultrametic starting tree:
tr_plastome <- multi2di(tr_plastome)
tr_plastome$edge.length[
  is.na(tr_plastome$edge.length) | tr_plastome$edge.length <= 0
] <- 1e-6

tr_plastome_ultra <- chronos(tr_plastome)
tr_plastome_ultra <- compute.brlen(tr_plastome_ultra, method = "Grafen")

write.tree(
  tr_plastome_ultra,
  here(
    "content/data/input/phylogenetic_tree/uapaca_start_ultrametric_plastome.tre"
  )
)
```

# IV. Convert alignment fastas into nexus format for RevBayes

The alignment fastas were copied from the original HybSuite folder in my Linux subsystem into the data/molecular_data folder. I chose the 10 genes with the most taxa represented in the alignment. 

For reference (only), here is the script I used to copy files from Linux to Windows.

```{bash, include=FALSE}
#| eval: false
#!/bin/bash

target_mol='/mnt/c/Users/frwo0304/Documents/GitHub/RevBayes-DEC-Uapaca/content/data/input/molecular_data'

source_mol='/home/frieda/HybSeq_pipeline_2026/data/output/06-Final_alignments/HRS'
source_plast='/home/frieda/Plastome_Phylogenomisc_2026/data/plastome_seq/plastome.fasta.concat.fasta'

mkdir -p $target_mol

cp \
"$source_mol/6295.aln.trimmed.fasta" \
"$source_mol/7029.aln.trimmed.fasta" \
"$source_mol/5594.aln.trimmed.fasta" \
"$source_mol/5620.aln.trimmed.fasta" \
"$source_mol/5922.aln.trimmed.fasta" \
"$source_mol/6016.aln.trimmed.fasta" \
"$source_mol/6620.aln.trimmed.fasta" \
"$source_mol/4527.aln.trimmed.fasta" \
"$source_mol/5454.aln.trimmed.fasta" \
"$source_plast/plastome_concat.fasta" \
"$target_mol/"

# check if they are there:
ls -1 "$target_mol"


# target folder:
target_tre='/mnt/c/Users/frwo0304/Documents/GitHub/RevBayes-DEC-Uapaca/content/data/input/phylogenetic_tree'
mkdir -p $target_tre


# source nuclear tree:
source_tre_nuc='/home/frieda/HybSeq_pipeline_2026/data/output/08-Coalescent_analysis/HRS/03-Species_tree/wASTRAL/wASTRAL_HybSuite_HRS.rr.tre'

# source plastome tree:
source_tre_plastome='/home/frieda/Plastome_Phylogenomisc_2026/data/plastome_seq/plastome.fasta.concat.treefile'

# copy nuclear tree
cp \
"$source_tre_nuc" \
"$target_tre/"

# copy plastome tree
cp \
"$source_tre_plastome" \
"$target_tre/"


# check if they are there:
ls -1 "$target_tre"

```

Reformat the alignments to the .nexus format needed in RevBayes from .fasta:

```{r}
#| eval: false
fasta_files <-
  list.files(
    "content/data/input/molecular_data",
    pattern = "\\.fasta$",
    full.names = TRUE
  )
out_files <-
  sub("\\.fasta$", ".nex", fasta_files)

for (i in seq_along(fasta_files)) {
  fasta_file <-
    fasta_files[i]

  out_file <-
    out_files[i]

  aln <-
    read.dna(
      fasta_file,
      format = "fasta",
      as.character = TRUE
    )

  ntax <-
    nrow(aln)

  nchar <-
    ncol(aln)

  taxa <-
    rownames(aln)

  seqs <-
    apply(
      aln,
      1,
      paste0,
      collapse = ""
    )

  name_width <-
    max(nchar(taxa)) + 12

  # Write new file:
  con <-
    file(out_file, open = "w")

  # Write header for each file:
  writeLines("#NEXUS", con)
  writeLines("begin data;", con)
  writeLines(paste0("dimensions ntax=", ntax, " nchar=", nchar, ";"), con)
  writeLines("format datatype=dna missing=? gap=-;", con)
  writeLines("matrix", con)

  for (i in seq_along(taxa)) {
    # Create line for each taxon with formatted name and sequence
    line <-
      paste0(
        format(taxa[i], width = name_width, justify = "left"),
        tolower(seqs[i])
      )
    writeLines(line, con)
  }

  writeLines(";", con)
  writeLines("end;", con)

  close(con)
}
```

```{bash}
#| eval: false

# rename the files
i=1

for f in *.aln.trimmed.nex; do
    mv "$f" "uapaca_gene_${i}.nex"
    ((i++))
done

```


# V. Download and set up RevBayes in project folder

Download RevBayes v1.3.2 for Linux directly from the GitHub releases page, and extract it to the /content folder.

```{r}
#| eval: false
url <- "https://github.com/revbayes/revbayes/releases/download/v1.3.2/revbayes-v1.3.2-linux64.tar.gz"
dest <- "content/RevBayes_Linux.tgz"

# 2. Download directly to the /content folder
download.file(url, destfile = dest, method = "libcurl")

# 3. Extract it using R's internal tar tool
utils::untar(dest, exdir = "content/")
```


## Starting RevBayes:

Start RevBayes in the terminal by starting the rb executable from the extracted folder.

```{bash}
#| eval: false

# to start RevBayes from the terminal, run this in bash:
./content/revbayes-v1.3.2/bin/rb

# run this in RevBayes terminal:
setwd("/mnt/c/Users/frwo0304/Documents/GitHub/RevBayes-DEC-Uapaca/")

# to run scripts from terminal, do: (for example)
source("./scripts/0_Setup_MCMC.Rev")
```

# References:

```{r}
sessionInfo()

citation("here")
citation("ape")
citation("dplyr")
citation("ggtree")
citation("ggplot2")
citation("RevGadgets")
citation("magick")
citation("BiocManager")
```